---
title: "Market Basket Analysis"
output: github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

# Libraries

```{r, message=FALSE}

library(readxl)
library(tidyverse)
library(lubridate)
library(arules)
library(arulesViz)
library(tidyquant)
library(plotly)

```

# Data

We will use the Online Retail data available at [UCI repository] (https://archive.ics.uci.edu/ml/datasets/online+retail).

```{r}

retails_df <- read_excel("Online Retail.xlsx")

head(retails_df,10) %>% knitr::kable()

```

# EDA

We will first have a look at the data before we do the analysis.

**Most Purchased Items**

```{r}

theme_set(theme_tq())

retails_df %>% 
  dplyr::count(Description, sort = TRUE) %>% 
  drop_na() %>% 
  head(15) %>% 
  ggplot(aes(x = reorder(Description,n), y = n))+
  geom_col(aes(fill = 'blue'), show.legend = FALSE)+
  coord_flip()+
  labs(title = "Top 15 Products Purchased",x = "", y = "Counts")+
  scale_fill_tq()

```


**Summary of Items purchased per transaction**

```{r}

retails_df %>% 
  count(InvoiceNo, sort = T) %>% 
  summarise_at(vars(n), funs(mean, median, min, max))


```

**Average itmes purchased per transaction by country**

```{r}

p1 <- retails_df %>% 
  count(InvoiceNo, Country) %>% 
  group_by(Country) %>% 
  summarise(total_trans = n(),
            total_items = sum(n)) %>% 
  mutate(avg_items_per_trans = total_items/total_trans,
         Country = fct_reorder(Country,avg_items_per_trans)) %>%
  ggplot(aes(x = Country,y = avg_items_per_trans))+
  geom_col(aes(fill = 'blue'), show.legend = FALSE)+
  coord_flip()+
  labs(title = "Average items purchased per transaction by Country",
       x = "", y = "Average items Purchased")+
  scale_fill_tq()

ggplotly(p1, tooltip = c("x","y")) %>% hide_legend()
 

```


**What time of the day most purchase take place..**

```{r}

p2 <- retails_df %>% 
  mutate(hr_day = hour(InvoiceDate)) %>% 
  group_by(InvoiceNo) %>% 
  slice(1) %>% 
  ungroup() %>% 
  mutate(hr_day = as.factor(hr_day)) %>% 
  ggplot(aes(x = hr_day))+
  geom_bar(aes(fill = 'blue'), show.legend = FALSE)+
  labs(title = "Distribution of Hour of Transaction", x = 'Hour of Day',
       y = "Transactions")+
  scale_fill_tq()

ggplotly(p2, tooltip = c("x","y")) %>% hide_legend()
 

```



# Preprocessing Data

Our data is currently in data frame format, but we need the data in 'transactions' format to perform market basket anaysis with arules package.

**Preparing data in the transactions format for association rules analysis**


```{r}
library(plyr)

retails_transactions <- ddply(.data = retails_df, 
                              .variables = c('InvoiceNo'),
                              .fun = function(df)paste(df$Description, collapse = ",")
                               )


retails_transactions %>% head() %>% knitr::kable()

```


We can see now the data is arranged in transaction format where each row represents a transaction. For further analysis we will only need the coulumn that contains the items purchased and each item in the coulmn represents a tansaction.

**Saving the transaction data**

```{r}

transactions <- retails_transactions %>% 
  dplyr::rename(items = V1) %>% 
  select(items) 

write.csv(transactions,"transactions.csv",quote = FALSE, row.names = FALSE )


```



**Read the transactions data that we just saved for rule mining**

```{r}

tr <- read.transactions("transactions.csv",format = 'basket',
                               sep = ",", rm.duplicates = TRUE, skip = 1)

tr

```

```{r}
summary(tr)

```

Inspect few of the transactions

```{r}
inspect(head(tr))
```


```{r}

itemFrequencyPlot(tr,topN = 10, type = "absolute", horiz = TRUE, col = "blue")
```


# Generating Rules

```{r}

rules <- apriori(tr, parameter = list(support = 0.009,  confidence = 0.7, 
                                             maxlen = 4) )

```

```{r}

summary(rules)

```

**Inspecting the rules**

```{r}

inspect(head(rules))

```


**Removing Redundant rules**

```{r}
# reduntand rules

inspect(rules[is.redundant(rules)])

```

```{r}
# Removing redundant rules

rules <- rules[!is.redundant(rules)]

```


**Top 10 Rules bsed on the lift**

```{r}

top_rules <- sort(rules, by = 'lift', decreasing = TRUE)[1:10]

inspect(top_rules)

```


**What are the products influencing sugar purchase?**

```{r}

sugar_rules <- apriori(tr, parameter = list(support = 0.009, confidence = 0.7),
                       appearance = list(default = 'lhs', rhs = 'SUGAR'))

```

```{r}

inspect(sort(sugar_rules, by = 'lift', decreasing = TRUE))

```

**What products are purchased with sugar?**

```{r}
with_sugar <- apriori(tr, parameter = list(support = 0.009, confidence = 0.7),
                       appearance = list(default = 'rhs', lhs = 'SUGAR'))

```

```{r}
 inspect(sort(with_sugar, by = 'lift', decreasing = TRUE))

```


# Visualization

**Scatter plot of all rules**

```{r}

plot(rules)

```


**Interactive Scatter plot of top 10 rules**

```{r}
plotly_arules(top_rules)
```

** Interactive Matrix Plot**

```{r}
plotly_arules(rules, method = "matrix")

```


**Network Plot for top 10 rules**

```{r}

plot(top_rules, method = 'graph', engine = 'htmlwidget'
       )

```


